\sysfull

### 2018-04-13 paper/paging.tex
alg 1 moved to top
added todos section
added some things to paging

\sys exposes non-volatile memory for \emph{protected} variables to
the application as a virtual address space partitioned into pages.
Paged virtual memory is a key enabler of task coalescing, which
we described in Section~\ref{sec:task_coalescing}. Virtualization simplifies
coalescing, because the system gets the freedom to move application's data
between volatile and non-volatile memory transparently.
%
\sys manages memory such that application tasks operate primarily from
efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
as we showed in Figure~\ref{fig:framEnergy}. \sys runtime implements memory
virtualization by moving pages of the address space from the large,
non-volatile memory to a small, volatile memory on demand and writing all dirty
pages back atomically. Since memory moves are at page granularity, \sys
exploits hardware\hyp{}accelerated block copies.

@@

As they are
accessed by a task, pages are swapped from their home location
in the direct\hyp{}mapped non-volatile memory, called
\texttt{\underline{store}} (underline denotes allocation in non-volatile
memory) into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
\texttt{working} buffer is allocated to take up the volatile memory that
remains after space for stack is allocated.
%
\sys uses two-phase commit to write dirty pages from the \texttt{working} buffer
back to their home location in \texttt{\underline{store}} atomically,
ensuring data consistency even with power failures.
%
When any dirty page is swapped out from the \texttt{working} buffer due to
eviction or task transition from last coalesced task, the first phase writes
the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
buffer, and the second phase writes it from
\texttt{\underline{shadow}} to \texttt{\underline{store}}.
%
Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
time between volatile and non-volatile memory, \sys's page-based
design moves contiguous blocks of data.
%
\sys takes full advantage of hardware\hyp{}accelerated block copies using
Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x
fewer cycles on MSP430FRxxxx.

@@

\begin{algorithm}
	\caption{\texttt{RVAR}(variable $v$)}
	\label{algo:rwar}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
		\State $t \leftarrow \Call{GetTag}{v}$ 
        \State $i \gets \{ j\ |\ \Call{GetTag}{\texttt{working}[j]} = t \}$ \Comment{Look in the page buffer}
		\If { $i = \emptyset$ } \Comment{Is the variable in a resident page?}
		\State	$i \gets \Call{PageFault}{t}$ \Comment{Page in from non-volatile memory}
		\EndIf
		\State $o \leftarrow \Call{GetOffset}{v}$ 		
		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return directly from page buffer}
	\end{algorithmic}
\end{algorithm}

@@

\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent
protected data. During task execution, modifications to protected data do not
affect their original memory locations, because a task reads and writes the
volatile \texttt{working} buffer only, and modified pages are kept in
\texttt{\underline{shadow}} until commit. A power failure erases the contents
of the page buffer, preventing a re-executing task from observing updates in
the page buffer from a previous execution attempt. Clearing \texttt{\underline{shadow}}
entries as part of the second phase commit (Line 13), ensures that all accesses to
protected variables in subsequent tasks correctly access their original memory
locations in \texttt{\underline{store}}.
%
%The non-volatile commit bit persists
%across power failures and \sys repeatedly attempts to commit until it succeeds,
%despite arbitrary power failures.

%\subsection{Managing Paging Overhead}
%
%The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations. To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops. We copied blocks of data of different size between memory regions on a WISP~\cite{wisp} using a hand-written loop and using DMA (refer to Section~\ref{sec:results_hardware} for hardware setup details). Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks of data like a page, hardware DMA is considerably better than software in both run time and energy efficiency. \sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
%
%\begin{figure}[t]
%	\centering
%	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
%	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
%	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
%	\label{fig:dmaTimeEnergy}
%\end{figure}

>>>>

\sys exposes non-volatile memory for \emph{protected} variables to
the application as a virtual address space partitioned into pages.
Paged virtual memory is a key enabler of task coalescing, which
we described in Section~\ref{sec:task_coalescing}. Virtualization simplifies
coalescing, because the system gets the freedom to move application's data
between volatile and non-volatile memory transparently.
%
\sys manages memory such that application tasks operate primarily from
efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
as we showed in Figure~\ref{fig:framEnergy}. \sys runtime implements memory
virtualization by moving pages of the address space from the large,
non-volatile memory to a small, volatile memory on demand and writing all dirty
pages back to non-volatile memory atomically. Since memory moves at page granularity, \sys
exploits hardware\hyp{}accelerated block copies using DMA.

@@

Paging moves data between non-volatile and volatile memory. A page has a home
location in non-volatile memory which we refer to as the page's {\tt store}.
As a task accesses pages, they are swapped from their home location
in the direct\hyp{}mapped non-volatile memory, called
\texttt{\underline{store}} (underlined locations are non-volatile) 
into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
\texttt{working} buffer is allocated to take up the volatile memory that
remains after space for stack is allocated.
%
\sys uses two-phase commit to write dirty pages from the \texttt{working} buffer
back to their home location in \texttt{\underline{store}} atomically,
ensuring data consistency even with power failures.
%
When any dirty page is swapped out from the \texttt{working} buffer due to
eviction or task transition from last coalesced task, the first phase writes
the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
buffer, and the second phase writes it from
\texttt{\underline{shadow}} to \texttt{\underline{store}}.
%
Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
time between volatile and non-volatile memory, \sys's page-based
design moves contiguous blocks of data.
%
\sys takes full advantage of hardware\hyp{}accelerated block copies using
Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x
fewer cycles on MSP430FRxxxx.

@@

\begin{algorithm}[t]
	\caption{\texttt{RVAR}(variable $v$)}
	\label{algo:rwar}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
		\State $t \leftarrow \Call{GetTag}{v}$ 
        \State $i \gets \{ j\ |\ \Call{GetTag}{\texttt{working}[j]} = t \}$ \Comment{Look in the page buffer}
		\If { $i = \emptyset$ } \Comment{Is the variable in a resident page?}
		\State	$i \gets \Call{PageFault}{t}$ \Comment{Page in from non-volatile memory}
		\EndIf
		\State $o \leftarrow \Call{GetOffset}{v}$ 		
		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return directly from page buffer}
	\end{algorithmic}
\end{algorithm}

@@

\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent
protected data. During task execution, modifications to protected data do not
affect their original memory locations, because a task reads and writes the
volatile \texttt{working} buffer only, and modified pages are kept in
\texttt{\underline{shadow}} until commit. A power failure erases the contents
of the page buffer, preventing a re-executing task from observing updates in
the page buffer from a previous execution attempt. Clearing \texttt{\underline{shadow}}
entries as part of the second phase commit (Line 13), ensures that all accesses to
protected variables in subsequent tasks correctly access their original memory
locations in \texttt{\underline{store}}.
%
%The non-volatile commit bit persists
%across power failures and \sys repeatedly attempts to commit until it succeeds,
%despite arbitrary power failures.
\subsection{Dynamic vs. Static Privatization}
\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
\TODO{pass by ref might not work the right way}
\begin{verbatim}

GLOBAL_(x)

foo(*y){ 
  *y++; 
}

TASK( ){ 
  ... ;  
  foo( &GV(x) ) ; 
}

\end{verbatim}
%\subsection{Managing Paging Overhead}
%
%The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations. To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops. We copied blocks of data of different size between memory regions on a WISP~\cite{wisp} using a hand-written loop and using DMA (refer to Section~\ref{sec:results_hardware} for hardware setup details). Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks of data like a page, hardware DMA is considerably better than software in both run time and energy efficiency. \sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
%
%\begin{figure}[t]
%	\centering
%	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
%	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
%	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
%	\label{fig:dmaTimeEnergy}
%\end{figure}
